---
title: "Project Reproductible Research Week 2"
author: "Thuy Linh"
date: "8/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) 

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as 𝙽𝙰) </br>
date: The date on which the measurement was taken in YYYY-MM-DD format </br>
interval: Identifier for the 5-minute interval in which measurement was taken </br>
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset. 

## 1. Code for reading in the dataset and/or processing the data 

Dowload, unzip and read data.
```{r}
fileUrl<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="./repdata-data-activity.zip")
unzip(zipfile="repdata-data-activity.zip")
data<-read.csv("./activity.csv")
```

## 2. Histogram of the total number of steps taken each day

Calculate the total number of steps taken per day by tapply function then create the histogram plot.

```{r}
hist(as.table(tapply(data$steps,data$date,sum)),xlab="steps",col="blue",main="Total number of steps taken each day")
```

## 3. Mean and median number of steps taken each day 

Compute the mean and the median of the steps each day using tapply function.

```{r}
tapply(data$steps,data$date,mean)
tapply(data$steps,data$date,FUN=median)
```

## 4. Time series plot of the average number of steps taken

Create the data frame meansteps contains the dates and the correspondence average number of steps then plot them.

```{r}
meansteps<-aggregate(list(Steps=data$steps),list(Date=data$date),mean)
meansteps$Date<-strptime(as.character(meansteps$Date),"%Y-%m-%d")
plot(meansteps$Date,meansteps$Steps,type="l",xlab="Date",ylab="Steps",main="Average number of steps taken each day")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

Create the data frame steps_interval contains the intervals and the correspondence average number of step, take the interval that have the biggest average step.

```{r}
steps_interval<-aggregate(list(steps=data$steps),list(interval=data$interval),FUN=mean,na.rm=TRUE)
steps_interval$interval[steps_interval$steps==max(steps_interval$steps)]

```

## 6. Code to describe and show a strategy for imputing missing data

Create a table counting the number of missing value, and the percentage of them

```{r}
table(is.na(data$steps))
prop.table(table(is.na(data$steps)))
```

We can probably impute the missing values by using the mean, the median, or the mode of the 5-minute intervals. I used the mean.

```{r}
newdata<-data
for (i in 1:nrow(newdata)){
  if (is.na(newdata$steps[i])==TRUE){
    newdata$steps[i]<-steps_interval$steps[steps_interval==data$interval[i]]
  }
}

```

## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
hist(as.table(tapply(newdata$steps,newdata$date,sum)),xlab="steps",col="blue",main="Total number of steps taken each day")
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Calculate the average number of steps taken per 5-minute interval for weekdays and weekends, then create the panel plot to compare them

```{r}
library(timeDate)
library(ggplot2)
weekday<-as.vector(tapply(newdata$steps[isWeekend(strptime(as.character(newdata$date),"%Y-%m-%d"),wday=1:5)==F],
         newdata$interval[isWeekend(strptime(as.character(newdata$date),"%Y-%m-%d"),wday=1:5)==F],mean))
weekend<-as.vector(tapply(newdata$steps[isWeekend(strptime(as.character(newdata$date),"%Y-%m-%d"),wday=1:5)==T],
         newdata$interval[isWeekend(strptime(as.character(newdata$date),"%Y-%m-%d"),wday=1:5)==T],mean))
weekdays<-as.data.frame(cbind(steps=c(weekday,weekend),interval=rep(unique(newdata$interval),2),
          weekday=c(rep("weekday",288),rep("weekend",288))))
qplot(as.numeric(interval),as.numeric(steps),data=weekdays,facets=weekday~.,geom="line",col=weekday,
      xlab="interval",ylab="steps",main="Average number of steps taken per 5-minute interval weekdays-weekends")
```

